---
title: "Week 1"
author: "Sam Gow"
format: revealjs
editor: visual
---

# Lecture content

Coding in week 1 lecture content is limited to introductory tasks including:

------------------------------------------------------------------------

### Basic mathematical operations

```{r basic_mathemetical_operations, echo = TRUE, include = TRUE}
# Addition
5+2
# Subtraction
6-8
#Multiplication
3*4
# Division
3/4
#Powers
3^4

```

------------------------------------------------------------------------

### Assignment and evaluation of variables

```{r assignment_and_evaluation, echo = TRUE, include = TRUE}
# Assignment
a = 3
b = 7
the.value = 12
the.index = 2
the.index_2 = 3
# Evaluation
a * b
the.value * a
a^the.index + b^the.index_2
```

------------------------------------------------------------------------

### Vectors and arrays

```{r vectors_arrays, echo = TRUE, include = TRUE}
# Creating vectors and arrays
c(3,2,1,0)
4:8
array(0,3)
array(1,10)
array(2,200)
array(c(1,2,3), 10)
array(c(4,5,6), 12)
array(0, c(3,4))
```

------------------------------------------------------------------------

### Manipulating vectors and arrays

```{r manipulating vectors_arrays, echo = TRUE, include = TRUE}
# Assignment
a = array(0,20)
a
# Updating
a[5] = 1
a[c(3,7,11)] = c(2,6,1)
a[17:20] = c(1,2,1,4)
a
# Combining 
cbind(c(1,2,3,7,9),c(21,2,1,5,6))
cbind(c(1,2,3,7,9),c(1,9,7,2,1),array(6,5))

rbind(c(3,6,1,92),c(10,3,1))
rbind(c(3,6,1,92),c(3,2,1,8))
rbind(c(3,6,1,92),c(4,1,12,1,2))
```

------------------------------------------------------------------------

### Basic operations on vectors

```{r operations_on_vectors, echo=TRUE}
# Combine vectors and normal numbers
c(1,2,4) + 4
c(1,2,3) * 3
c(1,2,3) ^ 2

#Operations between vectors
v1 = c(1,6,7,9)
v2 = c(-1,2,1,-2)

v1 + v2
v1 - v2
v1 * v2
v1/v2
v1 ^ v2 
```

------------------------------------------------------------------------

### Existing functions

```{r existing_functions, echo = TRUE}
# Assign the following vectors
a = c(1,8,3,9)
b = c(2,2,1,1)
d = c(3,4,6,81,9)

# Apply functions
sum(a)
prod(d)
length(b)
sum(a*b)
sum(a^b)
min(d)
max(a,d)

# Apply combined functions
prod(a)*(1/length(b))
min(max(a),max(b))
```

------------------------------------------------------------------------

# Practical

The following slides cover the content in the Week 1 practical pdf handout.

## Creating new functions

```{r creating_new_functions, echo = TRUE}
## Define our own mean function 
our_mean = function(x) {
  sum(x)/length(x)
}

our_mean(a)

our_mean(d)

## An alternate form of our mean function

our_mean_2 = function(x) {
    n <- length(x) # Calculate n
    s <- sum(x) # Calculate sum, s
    output <- s/n  #use, s and n to calcuate output 
    output  # Return output
}

our_mean_2(d)

```

------------------------------------------------------------------------

## An error example

The following function does not return a value as the final line of code involves an assignment.

```{r nothing_happens, echo = TRUE}

nothing_happens <- function(x) {
  sum(x) +2
  a <- prod(x)
}

```

------------------------------------------------------------------------

## Geometric and arithmetic mean functions.

Geometric mean is calculated by multiplying a set of integers and taking the nth power root where n is the length of the set.

::::: columns
::: {.column width="65%"}
```{r geometric_mean, echo = TRUE}

geom_mean = function(x) {
  if(length(x)==0){
    warning("geom_mean works with objects of length 1 or greater")
    return(NA_real_)}
  if(min(x)<0){
    warning("geom_mean works with positive numbers")
    return(NA_real_)}
  
  # Calculate power value to use - the length of x.
  power = length(x)
  # Calculate product of values of x,
  product = prod(x)
  product^(1/power)
}
```
:::

::: {.column width="35%"}
```{r geom_examples, echo = TRUE}

geom_mean(a)

geom_mean(
  x = c(1,2,3))

geom_mean(
  x = c())

geom_mean(
  x = c(1,-1))

```
:::
:::::

------------------------------------------------------------------------

Arithmetic mean is calculated by taking the sum of a set of numbers and dividing by the length of the set

::::: columns
::: {.column width="65%"}

```{r arithmetic_mean, echo = TRUE}

arithmetic_mean = function(x) {
   if(length(x)==0){
    warning(
      "arithmetic_mean works with objects of length 1 or greater")
    return(NA_real_)}
  n = length(x)
  s = sum(x)
  s/n
}

```

:::
::: {.column width="35%"}

``` {r arithmetic_examples, echo = TRUE}
arithmetic_mean(a)

arithmetic_mean(d)

arithmetic_mean(c())
```
:::
:::::
